---
title:"StochasticPracticals"
author:"VinaySanjayPatil"
date:"`rSys.Date()`"
output:
word_document:default
---

# CalculatinglimitingDistributionofMArkovChain
### 3.1ConsiderafiniteMarkovChain{Xn,n∈N}withstatespaceS={0,1,2,3}andTPMisgiven.FindthelimitingprobabilitiesofthisMarkovchain.

```{r}
library(expm)
P=matrix(c(2,1,1,0,
0,1,1,2,
1,2,1,0,
1,1,1,1),
ncol=4,byrow=TRUE)/4
#(a)Limitingdistributionvialinearequations
A=rbind(t(P)-diag(4),rep(1,4))
b=c(rep(0,4),1)
pi_eq=solve(t(A)%*%A,t(A)%*%b)
#(b)Powermethod
P50=P%^%50
P100=P%^%100
#(c)Differentinitialdistributions
alpha1=c(1,0,0,0)
alpha2=c(0.25,0.25,0.25,0.25)
alpha1%*%P100
alpha2%*%P100
pi_eq

```
### Conclusion:
#### 1.Thechainisergodic(irreducible+aperiodic),soitconvergestoauniquestationarydistribution.

#### 2.Thestationarydistributionis:π=(0.2292,0.3125,0.25,0.2083)

#### 3.Nomatterthestartingstate,aftermanytransitionstheprobabilityofbeingineachstatestabilizestothesevalues.

#### 4.ThefactthatP^100hasidenticalrowsequaltoπconfirmsconvergence.
#-----------------------------------------------------------------------------------------------------------------------------------------------

###3.2ConsideraMarkovchain{Xn,n∈N}withstatespaceS={0,1,2,3,4}.SupposeP04=1andsupposethatwhenthechainisinstatei,i>0,thenextstateisequallylikelytobeanyofthestates0,1,···,i−1.FindthelimitingprobabilitiesofthisMarkovchain.
```{r}
rm(list=ls())
library(expm)
P=matrix(c(0,0,0,0,1,
1,0,0,0,0,
1/2,1/2,0,0,0,
1/3,1/3,1/3,0,0,
1/4,1/4,1/4,1/4,0),ncol=5,byrow=TRUE)

#(b)Limitingdistribution
A=rbind(t(P)-diag(5),rep(1,5))
b=c(rep(0,5),1)
pi_eq=solve(t(A)%*%A,t(A)%*%b)

#(c)Powermethod
P100=P%^%100
#(d)DifferentInitialstates
alpha0=c(1,0,0,0,0)
alpha4=c(0,0,0,0,1)
alpha0%*%P100
alpha4%*%P100
pi_eq
```

# 4.RealizationofBienayme-Galton-WatsonBranchingProcess

### 4.1Let{Xn,n∈N}beaBienayme–Galton–WatsonbranchingprocesswithoffspringdistributionP(ξ=0)=0.2,P(ξ=1)=0.3,P(ξ=2)=0.2,P(ξ=3)=0.3.
#### IfX0=2,obtainrealizationsofX1,X2,...,X5.
#### (a)Generateonerealizationoftheprocess.
##### (b)Repeattherealizationusingadifferentrandomseed.
#### (c)Generate10independentrealizationsandobservethevariability.
#### (d)Identifyifextinctionoccursinanyrealization.
```{r}
rm(list=ls())
set.seed(1)
z=c(0,1,2,3)
pz=c(0.2,0.3,0.2,0.3)
#(a)Singlerealization
x=numeric(6)
x[1]=2
for(i in 1:5)
x[i+1]=sum(sample(z,size=x[i],prob=pz,replace=TRUE))
x
#(b)Differentseed
set.seed(10)
x2=numeric(6)
x2[1]=2
for(i in 1:5){
x2[i+1]=sum(sample(z,size=x2[i],prob=pz,replace=TRUE))
return(x2[i+1])}
x2
#(c)Multiplerealizations
paths=matrix(0,nrow=10,ncol=6)
for(j in 1:10){
paths[j,1]=2
for(i in 1:5){
paths[j,i+1]=sum(sample(z,size=paths[j,i],prob=pz,replace=TRUE))
}
}
paths

```

### 4.2
```{r}
rm(list=ls())
set.seed(2)
z=c(0,1,2,3)
pz=c(0.25,0.35,0.3,0.1)
x=numeric(9)
x[1]=3
for(i in 1:8)
x[i+1]=sum(sample(z,size=x[i],prob=pz,replace=TRUE))
x
```

### 4.3
```{r}
set.seed(3)
z=c(0,1,2,3,4,5)
pz=c(0.1,0.2,0.15,0.2,0.25,0.1)
paths=matrix(0,nrow=10,ncol=8)
for(j in 1:10){
paths[j,1]=1

for(i in 1:7){
paths[j,i+1]=sum(sample(z,size=paths[j,i],prob=pz,replace=TRUE))
}
}
paths
```
# 5. SimulationofPoissonProcess
### 5.1 Let{N(t),t≥0}beaPoissonprocesswithrateλ=x.
### (a)Simulatetheinterarrivaltimesandobtainthewaitingtimesuptot=y.
### (b)PlotthesamplepathofN(t)for0≤t≤y.
### (c)Repeatthesimulationusingadifferentrandomseedandcomparethepaths.
### Parameters:
### (i)x=2,y=5(ii)x=4,y=9.5(iii)x=9,y=7

```{r}
PoisProc=function(lambda,time){

inter=rexp(50,lambda)
arr=cumsum(inter)
arr=arr[arr<time]
n=length(arr)

t1=c(0,arr)
Nt=0:n

plot(t1,Nt,type="s",xlab="t",ylab="N(t)")
data.frame(t1,Nt)

}

set.seed(1)
PoisProc(2,5)

set.seed(2)
PoisProc(2,5)


```

### 5.2 SimulatethePoissonprocessatspecifiedtimepointswithoutsimulatinginterarrivaltimes.
(a)GenerateN(t)atthegiventimepoints.
(b)Repeatthesimulationandcompareoutputs.
(c)Comparethesimulatedvalueswithλt.
Parameters:
(a)t=1.5,2.2,3.8,7.5,8.8,λ=1
(b)t=1.23,2.21,2.83,6.05,7.08,17.8,λ=1.5

```{r}
rm(list = ls())
SimuPoiss = function(lambda,t)
{
  n = length(t)
  Nt = numeric(n)
  Nt[1] = rpois(1, lambda*t[1])
  for ( i in 2:n){
    Nt[i] = Nt[i-1] + rpois(1, lambda*(t[i] - t[i-1]))
    
    data.frame(t, Nt, lambda_t = lambda*t)
  }

}

set.seed(1)
SimuPoiss(1,c(1.5,2.2,3.8,7.5,8.8))

```

### 5.3 Given that N(x) = y, simulate the Poisson process over 0 ≤ t ≤ x.
### (a) Generate a conditional realization of the process.
### (b) Plot the resulting path.
### (c) Repeat the experiment and compare paths.
### Parameters:
### (i) x = 6, y = 5 (ii) x = 7, y = 10 (iii) x = 9, y = 7

```{r}
rm(list = ls())
SimPoiss = function(n, t)
{
  arr = sort(runif(n,0,t))
  t1 = c (0 , arr )
  Nt = 0:n
  plot ( t1 , Nt , type="s" , xlab="t" ,     ylab="N( t ) " )
  data.frame(t1,Nt)
}
set.seed(1)
SimPoiss(5 ,6)
  
```

